### 2021-04-30 submission.tex
" added previous submission work

%%
%% This is file `sample-acmtog.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmtog')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmtog.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[acmtog]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This infrmation is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2021}
\acmYear{2021}
%%\acmDOI{10.1145/1122445.1122456}

%%
%% These commands are for a JOURNAL article.
%%\acmJournal{TOG}
%%\acmVolume{37}
%%\acmNumber{4}
%%\acmArticle{111}
%%\acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Importance of Affective States in Privacy Calculus}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Thibault Irissou}
\affiliation{%
  \institution{Oberlin College}
  \city{Oberlin}
  \country{USA}}
\email{tirissou@oberlin.edu}

\author{Tara Bobinac}
\affiliation{%
  \institution{Oberlin College}
  \city{Oberlin}
  \country{USA}}
\email{tbobinac@oberlin.edu}

\author{Julia Cornell}
\affiliation{%
  \institution{Oberlin College}
  \city{Oberlin}
  \country{USA}}
\email{jcornel2@oberlin.edu}

\author{Riley Chinburg}
\affiliation{%
  \institution{Oberlin College}
  \city{Oberlin}
  \country{USA}}
\email{rchinbur@oberlin.edu}

\graphicspath{ {../images/} }

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Irissou, Bobinac, Chinburg, and Cornell}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
    Research historically understates the importance of affective states within privacy calculus frameworks. However, positive and negative affective states are critical in shaping human behavior. We aim to explore whether positive or negative affect more strongly reinforces privacy conscious behavior.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
%%\begin{CCSXML}
%%<ccs2012>
%% <concept>
%%  <concept_id>10010520.10010553.10010562</concept_id>
%%  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
%%  <concept_significance>500</concept_significance>
%% </concept>
%% <concept>
%%  <concept_id>10010520.10010575.10010755</concept_id>
%%  <concept_desc>Computer systems organization~Redundancy</concept_desc>
%%  <concept_significance>300</concept_significance>
%% </concept>
%% <concept>
%%  <concept_id>10010520.10010553.10010554</concept_id>
%%  <concept_desc>Computer systems organization~Robotics</concept_desc>
%%  <concept_significance>100</concept_significance>
%% </concept>
%% <concept>
%%  <concept_id>10003033.10003083.10003095</concept_id>
%%  <concept_desc>Networks~Network reliability</concept_desc>
%%  <concept_significance>100</concept_significance>
%% </concept>
%%</ccs2012>
%%\end{CCSXML}
%%
%%\ccsdesc[500]{Computer systems organization~Embedded systems}
%%\ccsdesc[300]{Computer systems organization~Redundancy}
%%\ccsdesc{Computer systems organization~Robotics}
%%\ccsdesc[100]{Networks~Network reliability}
%%

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
%%\keywords{datasets, neural networks, gaze detection, text tagging}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

Privacy literacy does not necessarily translate into privacy conscious behavior — the great privacy paradox. To describe this seemingly irrational decision-making process, many studies have explored different models of this privacy calculus \cite{Kehr2015}\cite{Jian2013}. These privacy-infringing decisions despite purported privacy concerns may result in negative affective states such as guilt and shame. \\

Question 1: To what extent are guilt or shame the result of privacy infringing decision-making? Is there a correlation between the privacy literacy of an individual and the amount of guilt/shame?

Hypothesis: We believe that guilt and shame will be felt in these situations. We hypothesize that the privacy literacy of an individual will positively correlate with the extent of these negative affective states. \\

Question 2: Does engaging in privacy-preserving behavior result in positive affective states?

Hypothesis 2: We predict that such affectives states will not correlate that strongly with privacy preserving behavior. \\

Question 3: Do these affectives states motivate people to adopt more privacy-preserving behavior?

Hypothesis 3: We predict that both negative affective states are not effective and that positive affective states are effective at motivating such behavior.

\section{Related Works}
\subsection{Privacy paradox, privacy calculus and affects}
The privacy paradox’s existence is for certain, as Barth and de Jong establish in their 2017 literature review. Contextual and environmental circumstances impact the rationality of decision-making, especially online. Studies demonstrate that irrationality in decisions-making regarding privacy is not necessarily circumvented through privacy literacy. \cite{Barth2017}

Some research took the direction of looking at privacy-preserving behavior as being the irrational decision [Cohen]. However, much recent literature assumes that decisions being made are rational insomuch as there is a level of privacy calculus that individuals perform when assessing the risks and rewards of disclosing information or violating their privacy \cite{Kehr2015}\cite{Jian2013}. Affect and affective states (emotions and moods) have been included in recent models of privacy calculus. These frameworks provide the basis for how we conceptualize affective impact on decision making regarding privacy.

\subsection{Negative affective states}
Teroni and Deonna explore in their paper the differentiation of shame and guilt. Their distinctions are the foundation for how we conceptualize negative affective states. We use the questionnaire as detailed in \cite{Rusch2007} as a model for how we build our survey.
\subsection{Positive affective states and Privacy}
Research places importance on the perceived social benefits that come from self-disclosure of information online in synchronous interactive environments\cite{Jian2013}. Such perception of benefits generates positive affective states from such divulsion of information.

In line with the investigation of positive affective states as a motivating factor in behavior shaping \cite{Taneja2014}, we are interested in how positive affective states as a result to making privacy-conscious and privacy-preserving decisions correlate with frequency of such decision-making.

\section{Methodology}
This is a mixed methods study. Surveys are sent out first to find correlations between privacy preserving behavior and the affective states that reinforce them to varying success. Interviews will be conducted with respondents from the survey who want to participate.
\pagebreak
\subsection{Questionnaires}
\nopagebreak The survey is designed to answer:
\begin{itemize}
    \item What privacy protecting behavior respondents actively engage in (from a list of common ones),
    \item How regular and consistent this behavior is
    \item The extent to which respondents feel good or bad when they engage in privacy-protective behavior
    \item The extent to which respondents feel good or bad when they do not engage in privacy-protective behavior
\end{itemize}

\noindent The survey was modeled after multiple established surveys:
\begin{itemize}
    \item Surveys to gauge privacy perception and behavior
    \item Use established questionnaires for measuring negative affects \cite{Rusch2007}
    \item Design questions with best practices for digital methodology in mind \cite{Hampton2017}
\end{itemize}

Quantitative data is analyzed to find correlations between behavior and affect.
\subsection{Interviews}
Findings in the quantitative data will be used by interviewers to help guide interviewees through semi-structured interviews. The objective of the interviews is to \textbf{a)} collect detailed descriptions of the affective states that are experienced before, during, and after engaging in privacy-related behavior, and \textbf{b)} explore the extent to which privacy-related behavior is shaped by the affective states surrounding them. Prior to addressing these, the interviewer will establish what the interviewee's privacy-related habits are and the extent of their knowledge about privacy. \\

The main guiding questions for the interviews are as follow:\\

\textbf{QUESTION 1}: How private and confidential is your information when you’re online? Are there certain activities that are riskier/safer than others?

\textbf{QUESTION 2}: Do you take any protective measures when using the internet? Why or why not?

\textbf{QUESTION 3}: How does the way that you currently browse the internet make you feel in regards to how secure you believe your information to be?

\section{Findings}
\subsection{Survey}
67 responses were collected from people over the age of 18. Respondents were primarily  college students, of which a majority are enrolled at Oberlin College.

\includegraphics[width=0.5\textwidth]{concerns}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{biblio}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\end{document}
\endinput
%%
%% End of file `sample-acmtog.tex'.